---
title: "cleaning"
output: html_document
date: "2024-04-26"
---

## Load libraries
```{r}
# Load the necessary library
library(dplyr)
library(broom)
library(car)
library(readr)
library(lubridate)
library(data.table)
library(MASS)
library(glmnet)
library(sandwich)
library(lmtest)
library(nortest)
library(boot)
library(ggplot2)
```

```{r}
# Load the datasets
aff <- read.csv("Affinity - State - Daily.csv")

aff <- aff %>%
  filter(!is.na(spend_all) & spend_all != ".")

aff$spend_all <- as.numeric(aff$spend_all)

aff <- aff %>%
  mutate(date = as.Date(paste(year, month, day, sep="-"), format="%Y-%m-%d"))

# Originally 2020-1-13 to 2024-4-7
aff <- aff %>%
  filter(date >= as.Date("2020-01-17") & date <= as.Date("2022-01-16"))
```

```{r}
# Load the datasets
emp <- read.csv("Employment - State - Weekly.csv")

emp <- emp %>%
  filter(!is.na(emp) & emp != ".")

emp$emp <- as.numeric(emp$emp)

emp <- emp %>%
  mutate(date = make_date(year, month, day_endofweek))

# Originally 2020-1-17 to 2024-2-16
emp <- emp %>%
  filter(date >= as.Date("2020-01-12") & date <= as.Date("2022-01-16"))
```

```{r}
# Load the datasets
small_biz <- read.csv("Womply - State - Weekly.csv")

small_biz <- small_biz %>%
  filter(!is.na(revenue_all) & revenue_all != ".")

small_biz$revenue_all <- as.numeric(small_biz$revenue_all)

small_biz <- small_biz %>%
  mutate(date = make_date(year, month, day_endofweek))

# Originally 2020-1-12 to 2022-2-6
small_biz <- small_biz %>%
  filter(date >= as.Date("2020-01-12") & date <= as.Date("2022-01-16"))
```

```{r}
# Load the datasets
math <- read.csv("Zearn - State - Weekly.csv")

math <- math %>%
  filter(!is.na(engagement) & engagement != ".")

math$engagement <- as.numeric(math$engagement)

math <- math %>%
  mutate(date = make_date(year, month, day_endofweek))

# Originally 2019-1-13 to 2024-4-21
math <- math %>%
  filter(date >= as.Date("2020-01-12") & date <= as.Date("2022-01-16"))
```

```{r}
# Load the datasets
covid <- read.csv("COVID - State - Daily.csv")

covid <- covid %>%
  filter(!is.na(case_rate) & case_rate != ".")

covid$case_rate <- as.numeric(covid$case_rate)

covid <- covid %>%
  mutate(date = as.Date(paste(year, month, day, sep="-"), format="%Y-%m-%d"))

# Originally 2020-1-21 to 2023-5-10
covid <- covid %>%
  filter(date >= as.Date("2020-01-12") & date <= as.Date("2022-01-16"))
```

# Define States
```{r}
state_abbreviations <- setNames(
  c("1"="AL", "2"="AK", "4"="AZ", "5"="AR", "6"="CA", "8"="CO", "9"="CT", "10"="DE", "11"="DC", "12"="FL", "13"="GA", 
    "15"="HI", "16"="ID", "17"="IL", "18"="IN", "19"="IA", "20"="KS", "21"="KY", "22"="LA", "23"="ME", "24"="MD", "25"="MA", 
    "26"="MI", "27"="MN", "28"="MS", "29"="MO", "30"="MT", "31"="NE", "32"="NV", "33"="NH", "34"="NJ", "35"="NM", "36"="NY", 
    "37"="NC", "38"="ND", "39"="OH", "40"="OK", "41"="OR", "42"="PA", "44"="RI", "45"="SC", "46"="SD", "47"="TN", "48"="TX", 
    "49"="UT", "50"="VT", "51"="VA", "53"="WA", "54"="WV", "55"="WI", "56"="WY"),
  c("1", "2", "4", "5", "6", "8", "9", "10", "11", "12", "13", 
    "15", "16", "17", "18", "19", "20", "21", "22", "23", "24", "25", 
    "26", "27", "28", "29", "30", "31", "32", "33", "34", "35", "36", 
    "37", "38", "39", "40", "41", "42", "44", "45", "46", "47", "48", 
    "49", "50", "51", "53", "54", "55", "56")
)

# Define region categories
northeast_states <- c('CT', 'ME', 'MA', 'NH', 'RI', 'VT', 'NY', 'NJ', 'PA')
west_states <- c('WA', 'OR', 'CA', 'NV', 'ID', 'MT', 'WY', 'UT', 'CO', 'NM', 'AZ', 'AK', 'HI')
south_states <- c('TX', 'OK', 'AR', 'LA', 'MS', 'AL', 'TN', 'KY', 'FL', 'GA', 'SC', 'NC', 'VA', 'WV', 'MD', 'DE', 'DC')
midwest_states <- c('ND', 'SD', 'NE', 'KS', 'MN', 'IA', 'MO', 'WI', 'IL', 'IN', 'OH', 'MI')

# Add a new column with state abbreviations using the lookup vector
emp <- emp %>%
  mutate(state_abbr = state_abbreviations[as.character(statefips)])

aff <- aff %>%
  mutate(state_abbr = state_abbreviations[as.character(statefips)])

# Create dummy variables for each region
aff <- aff %>%
  mutate(
    NE = ifelse(state_abbr %in% northeast_states, 1, 0),
    W = ifelse(state_abbr %in% west_states, 1, 0),
    S = ifelse(state_abbr %in% south_states, 1, 0),
    MW = ifelse(state_abbr %in% midwest_states, 1, 0)
  )

emp <- emp %>%
  mutate(
    NE = ifelse(state_abbr %in% northeast_states, 1, 0),
    W = ifelse(state_abbr %in% west_states, 1, 0),
    S = ifelse(state_abbr %in% south_states, 1, 0),
    MW = ifelse(state_abbr %in% midwest_states, 1, 0)
  )
```

# create wave dummies
```{r}
aff <- aff %>%
  mutate(
    wave1 = ifelse(date >= as.Date("2020-01-17") & date <= as.Date("2020-12-19"), 1, 0),
    wave2 = ifelse(date >= as.Date("2020-12-20"), 1, 0),
  )
```

# create wave dummies
```{r}
emp <- emp %>%
  mutate(
    wave1 = ifelse(date >= as.Date("2020-01-17") & date < as.Date("2020-12-25"), 1, 0),
    wave2 = ifelse(date > as.Date("2020-12-25"), 1, 0)
  )
```

```{r}
colnames(aff)
```

```{r}
colnames(emp)
```

## Merging emp and aff (5250*59)

```{r, warning=FALSE}
aff <- aff %>%
  mutate(across(matches("^spend_|^freq$"), ~ as.numeric(as.character(.))))

# Convert data frames to data.tables
setDT(aff)
setDT(emp)

# Convert dates to Date objects
aff[, date := as.Date(date)]
emp[, date := as.Date(paste(year, month, day_endofweek, sep = "-"))]

# Identify numeric columns for the mean calculation
numeric_cols <- names(aff)[sapply(aff, is.numeric)]

# Aggregate the daily data to weekly, calculating means only for numeric columns
aff_weekly <- aff[, lapply(.SD, mean, na.rm = TRUE), by = .(statefips, date = floor_date(date, "week")), .SDcols = numeric_cols]

aff_weekly$date <- aff_weekly$date + 5

# Now merge the weekly datasets
emp_aff <- emp[aff_weekly, on = .(date, statefips)]

emp_aff <- emp_aff[!is.na(emp_aff$year), ]
```

```{r}
colnames(emp_aff)
```

## Merging emp_aff with math (5311*78)

```{r}
emp_aff$date <- as.Date(emp_aff$date)
math$date <- as.Date(math$date)
math$date <- math$date - 2

emp_aff_math <- merge(emp_aff, math, by = c("date", "statefips"), all = TRUE)
```

```{r}
colnames(emp_aff_math)
```

## Merging emp_aff_math with small_biz (5352*93)

```{r}
emp_aff_math$date <- as.Date(emp_aff_math$date)
small_biz$date <- as.Date(small_biz$date)
small_biz$date <- small_biz$date + 5

emp_aff_math_biz <- merge(emp_aff_math, small_biz, by = c("date", "statefips"), all = TRUE)

emp_aff_math_biz <- emp_aff_math_biz %>%
  filter(date <= as.Date("2022-01-14"))
```

```{r}
colnames(emp_aff_math_biz)
```

## Ensuring all required columns are numeric to we can run lm

```{r}
emp_aff_math_biz <- emp_aff_math_biz %>%
  filter(!is.na(revenue_all))
emp_aff_math_biz$revenue_all <- as.numeric(emp_aff_math_biz$revenue_all)
emp_aff_math_biz$merchants_all <- as.numeric(emp_aff_math_biz$merchants_all)
emp_aff_math_biz$spend_all <- as.numeric(emp_aff_math_biz$spend_all)
emp_aff_math_biz$emp <- as.numeric(emp_aff_math_biz$emp)
emp_aff_math_biz$engagement <- as.numeric(emp_aff_math_biz$engagement)
```

## Merging emp_aff_math_biz with covid (5069*109)

```{r}
# Convert data frames to data.tables
setDT(covid)

# Identify numeric columns for the mean calculation
numeric_cols <- c("year", "month", "day", "statefips", "case_count", "new_case_count", "death_count", "new_death_count", "case_rate", "new_case_rate", "death_rate", "new_death_rate", "vaccine_count", "new_vaccine_count", "vaccine_rate", "new_vaccine_rate")
covid[, (numeric_cols) := lapply(.SD, as.numeric), .SDcols = numeric_cols]

# Aggregate the daily data to weekly, calculating means
covid_weekly <- covid[, lapply(.SD, mean, na.rm = TRUE), by = .(statefips, date = floor_date(date, "week")), .SDcols = numeric_cols]

covid_weekly$date <- covid_weekly$date - 2

# Now merge the weekly datasets
emp_aff_math_biz_covid <- emp_aff_math_biz[covid_weekly, on = .(date, statefips)]
```

```{r}
colnames(emp_aff_math_biz_covid)
```

## Either emp_aff_math_biz could be the final dataset (preffered I guess) - But covid waves here are no longer significant in the model

## Or emp_aff_math_biz_covid could be the final dataset - Since we have covid metrics but covid dataset does not have data for all state for the whole time period (5000 rows only)

## Merging emp_aff with small_biz for FWL 

```{r}
emp_aff_biz <- merge(emp_aff, small_biz, by = c("date", "statefips"), all = TRUE)

emp_aff_biz <- emp_aff_biz %>%
  filter(!is.na(revenue_all) & revenue_all != ".")

emp_aff_biz <- emp_aff_biz %>%
  filter(!is.na(emp) & emp != ".")

emp_aff_biz <- emp_aff_biz %>%
  filter(!is.na(spend_all) & spend_all != ".")
```

```{r}
cleaned_data <- na.omit(emp_aff_math_biz[, c("revenue_all", "wave1", "emp", "spend_all", "engagement", "merchants_all", "NE", "W", "S")])
```

