---
title: "textbook"
output: html_document
date: "2024-04-27"
---

## Univariate Regression 

#Confidence and Prediction Intervals

```{r}
model1 <- lm(spend_all ~ emp, data = emp_aff)
summary(model1)
```
```{r}
# Calculate the number of rows in your dataset
n <- nrow(emp_aff)

# Generate a random sample of indices for training data
train_indices <- sample(1:n, size = floor(0.8 * n), replace = FALSE)

# Create training data
training_data <- emp_aff[train_indices, ]

# Create testing data
testing_data <- emp_aff[-train_indices, ]  # Exclude the training indices

cat("Training data rows: ", nrow(training_data), "\n")
cat("Testing data rows: ", nrow(testing_data), "\n")
cat("Total rows: ", nrow(emp_aff), "\n")
```

```{r}
# Fit the model on the training data
model1 <- lm(spend_all ~ emp, data = training_data)

# Make predictions and calculate intervals on the testing data
ci_predictions <- predict(model1, newdata = testing_data, interval = "confidence")
pi_predictions <- predict(model1, newdata = testing_data, interval = "prediction")
```

```{r}
actual_vs_predicted <- data.frame(
  emp = testing_data$emp,
  Actual = testing_data$spend_all,
  Predicted = pi_predictions[, "fit"]
)

# Create the plot
ggplot(data = actual_vs_predicted, aes(x = emp)) +
  geom_line(aes(y = Actual), color = "grey40", alpha = 0.5) +  # Actual spend_all
  geom_line(aes(y = Predicted), color = "red") +  # Predicted spend_all
  ylab("Spend All") +
  ggtitle("Actual vs. Predicted Spend All") +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), plot.title = element_text(hjust = 0.5))
```
```{r}
predictions <- data.frame(
  emp = testing_data$emp,
  spend_all = testing_data$spend_all,
  Predicted = pi_predictions[, "fit"],
  Lower_PI = pi_predictions[, "lwr"],
  Upper_PI = pi_predictions[, "upr"],
  Lower_CI = ci_predictions[, "lwr"],
  Upper_CI = ci_predictions[, "upr"]
)

# Base plot with actual spend_all values
p <- ggplot(predictions, aes(x = emp)) +
  geom_point(aes(y = spend_all), color = "grey", alpha = 0.5, size = 0.6) +
  geom_line(aes(y = Predicted), color = "red", size = 1) +
  ylab("Spend All") +
  xlab("Emp") +
  ggtitle("Actual vs. Predicted Spend All") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

# Add the confidence interval as a ribbon
p <- p + geom_ribbon(aes(ymin = Lower_CI, ymax = Upper_CI), fill = "blue", alpha = 0.2)

# Add the prediction interval as a ribbon
p <- p + geom_ribbon(aes(ymin = Lower_PI, ymax = Upper_PI), fill = "green", alpha = 0.1)

# Show the plot
print(p)
```
## Could run the above for multivariate but it gets very complicated and he does not cover it either so we can just introduce with this and leave it at that 

Based on the graph titled "Actual vs. Predicted Spend All," which plots spend_all against emp, here's an interpretation:

Scatter Plot: The gray dots represent the actual data points of consumer spending (spend_all) for corresponding levels of employment (emp). This scatter distribution allows us to visualize the relationship between the two variables.
Red Line (Predicted Line): This line is the result of the univariate regression analysis, representing the best fit line through the data according to the linear model. It suggests that there is a positive relationship between employment levels and consumer spending, meaning as employment (emp) increases, consumer spending (spend_all) also tends to increase.
Green Bands (Prediction Intervals): The shaded green areas around the red line represent the prediction intervals. This interval is the range within which we expect future individual data points to fall, with a certain level of confidence (typically 95%). The fact that most of the actual data points (gray dots) fall within this green band indicates that the prediction intervals are capturing the variability of the actual data reasonably well.
Storyline and Implications:
If this graph represents a part of an economic analysis, it illustrates that employment levels are a significant predictor of consumer spending. The upward slope of the red line suggests that higher employment may lead to increased spending. This could be interpreted as a sign of economic health, where a robust job market fuels consumer confidence and spending habits.

However, it is important to note that while the positive relationship seems evident, the causality cannot be determined from this graph alone. Moreover, the presence of data points outside the prediction intervals is normal and expected, as previously explained. These could be anomalies or could represent other influencing factors not included in the model.

It's also important to consider that this is a simple linear regression model with only one predictor. In real-world scenarios, consumer spending is influenced by multiple factors, such as income levels, consumer confidence, credit availability, inflation, and more. To fully understand consumer spending patterns, a multivariate analysis with more predictors might be necessary.

Lastly, the prediction intervals indicate the model's uncertainty in predictions for individual observations. If the intervals are wide, it may imply high variability in consumer spending that's not explained by employment alone, suggesting the need for additional explanatory variables in the model.






