---
title: "employment models"
output: html_document
date: "2024-04-23"
---

## Modelling

# Need to check assumptions - check
# Need to add robust SE - check
# Need to check interactions + need to add other covariates (maybe remove state??)
# Need to try other transformations - tired log+1
# Need to check which maximises adjusted R squared : accordingly choose best model

```{r}
full_model1 <- lm(revenue_all ~ wave1 + emp + spend_all + engagement + merchants_all + NE + W + S, data = emp_aff_math_biz)
robust_se <- vcovHC(full_model1, type = "HC1")
coeftest(full_model1, robust_se) #hetero
summary(full_model1) #homo
plot(full_model1)
```

# Some variability so may need to try transformation but overall, fine (heteroskedasticity)

```{r}
ols.fit.hc0 = sqrt ( diag ( hccm ( full_model1 , type = "hc0" )))
ols.fit.hc1 = sqrt ( diag ( hccm ( full_model1 , type = "hc1" )))
ols.fit.hc2 = sqrt ( diag ( hccm ( full_model1 , type = "hc2" )))
ols.fit.hc3 = sqrt ( diag ( hccm ( full_model1 , type = "hc3" )))
ols.fit.hc4 = sqrt ( diag ( hccm ( full_model1 , type = "hc4" )))
full_model.coef = summary ( full_model1 )$ coef
tvalues = full_model.coef [ ,1]/ cbind ( full_model.coef [,2], ols.fit.hc0, ols.fit.hc1, ols.fit.hc2, ols.fit.hc3, ols.fit.hc4)
colnames ( tvalues ) = c(" ols ", "hc0", "hc1", "hc2", "hc3", "hc4")
round ( tvalues , 2)
```

# Test for multicollinearity - passed

```{r}
vif_values <- vif(full_model1)
print(vif_values)
```

# Test for normality

```{r}
# Perform Lilliefors (Kolmogorov-Smirnov) normality test
lillie_test <- lillie.test(residuals(full_model1))

# Print the results
print(lillie_test) #small p-value means not normal
```

# Test for heteroskedasticty and solution (GLM)

```{r}
# Perform Breusch-Pagan test
bp_test <- bptest(full_model1)
bp_value <- bp_test$p.value
print(bp_value) #small p-value means not homoskedastic
```

# Alternatively, you can use GLM

```{r}
glm_model <- glm(revenue_all ~ wave1 + emp + spend_all + engagement + merchants_all + NE + W + S, data = emp_aff_math_biz)
summary(glm_model)
```

```{r}
full_model2 <- lm(log(revenue_all +1) ~ wave1 + emp + spend_all + engagement + merchants_all + NE + W + S, data = emp_aff_math_biz)
robust_se <- vcovHC(full_model2, type = "HC1")
coeftest(full_model2, robust_se) #hetero
summary(full_model2) #homo
plot(full_model2)
```

```{r}
ols.fit.hc0 = sqrt ( diag ( hccm ( full_model2 , type = "hc0" )))
ols.fit.hc1 = sqrt ( diag ( hccm ( full_model2 , type = "hc1" )))
ols.fit.hc2 = sqrt ( diag ( hccm ( full_model2 , type = "hc2" )))
ols.fit.hc3 = sqrt ( diag ( hccm ( full_model2 , type = "hc3" )))
ols.fit.hc4 = sqrt ( diag ( hccm ( full_model2 , type = "hc4" )))
full_model.coef = summary ( full_model2 )$ coef
tvalues = full_model.coef [ ,1]/ cbind ( full_model.coef [,2], ols.fit.hc0, ols.fit.hc1, ols.fit.hc2, ols.fit.hc3, ols.fit.hc4)
colnames ( tvalues ) = c(" ols ", "hc0", "hc1", "hc2", "hc3", "hc4")
round ( tvalues , 2)
```

```{r}
vif_values <- vif(full_model2)
print(vif_values)
```


