---
title: "Models"
output: pdf_document
date: "2024-04-23"
---

# Need to check assumptions, which maximises adjusted R squared
# Need to add robust SE
# Need to check interactions + need to add other covariates
# Need to try other transformations

```{r}
total_spend_model <- lm(spend_all ~ wave1 + wave2 + wave3, data = aff)
summary(total_spend_model)
plot(total_spend_model)
```

```{r}
total_spend_model_q1 <- lm(spend_all_q1 ~ wave1 + wave2 + wave3, data = aff_clean)
summary(total_spend_model_q1)
plot(total_spend_model_q1)

total_spend_model_q2 <- lm(spend_all_q2 ~ wave1 + wave2 + wave3, data = aff_clean)
summary(total_spend_model_q2)
plot(total_spend_model_q2)

total_spend_model_q3 <- lm(spend_all_q3 ~ wave1 + wave2 + wave3, data = aff_clean)
summary(total_spend_model_q3)
plot(total_spend_model_q3)

total_spend_model_q4 <- lm(spend_all_q4 ~ wave1 + wave2 + wave3, data = aff_clean)
summary(total_spend_model_q4)
plot(total_spend_model_q4)
```

```{r}
aff$statefips <- as.factor(aff$statefips)

# Fit a regression model with statefips as a factor
state_model <- lm(spend_all ~ wave1 + wave2 + wave3 + statefips, data = aff)

# Print the summary of the model
summary(state_model)
plot(state_model)
```

```{r}
# List of spending categories
spending_vars <- c("spend_aap", "spend_acf", "spend_aer", "spend_apg", 
                   "spend_durables", "spend_nondurables", "spend_grf", 
                   "spend_gen", "spend_hic", "spend_hcs", "spend_inperson", 
                   "spend_inpersonmisc", "spend_remoteservices", "spend_sgh", 
                   "spend_tws", "spend_retail_w_grocery", "spend_retail_no_grocery", 
                   "spend_all_incmiddle", "spend_all_q1", "spend_all_q2", 
                   "spend_all_q3", "spend_all_q4")

# Loop through each spending variable and fit a model
for (spend_var in spending_vars) {
  # Fit a linear model for each spending category
  model <- lm(formula = paste(spend_var, "~ wave1 + wave2 + wave3 + statefips"), data = aff)
  
  # Use broom to tidy the model summary
  tidy_model <- tidy(model)
  
  # Print the tidy model summary for each spending category
  cat("\nCoefficient Table for", spend_var, ":\n")
  print(tidy_model)
}
```

# Need to check assumptions, which maximises adjusted R squared
# Need to add robust SE
# Need to check interactions + need to add other covariates
# Need to try other transformations

```{r}
total_emp_model <- lm(emp ~ wave1 + NE + W + S, data = emp)
summary(total_emp_model)
plot(total_emp_model)
```

```{r}
# Calculate VIF values
vif_values <- vif(total_spend_model)
print(vif_values)
```

```{r}
# Calculate VIF values
vif_values <- vif(total_emp_model)
print(vif_values)
```

# Need to check assumptions
# Need to add robust SE
# Need to check interactions + need to add other covariates
# Need to try other transformations
# Need to check which maximises adjusted R squared : accordingly choose best model

```{r}
emp_aff$spend_all <- as.numeric(emp_aff$spend_all)
emp_aff$emp <- as.numeric(emp_aff$emp)

model1 <- lm(spend_all ~ wave1 + emp + NE + W + S, data = emp_aff)
summary(model1)
plot(model1)

model2 <- lm(log(spend_all + 1) ~ wave1 + emp + NE + W + S, data = emp_aff)
summary(model2)
plot(model2)

model3 <- lm(log(spend_all + 1) ~ wave1 + log(emp + 1) + NE + W + S, data = emp_aff)
summary(model3)
plot(model3)
```

```{r}
# Calculate VIF values
vif_values1 <- vif(model1)
print(vif_values1)
vif_values2 <- vif(model2)
print(vif_values2)
vif_values3 <- vif(model3)
print(vif_values3)
```

# Model with emp_aff_math_biz_covid

```{r}
model4 <- lm(revenue_all ~ wave1 + emp + spend_all + engagement + merchants_all + NE + W + S + case_rate, data = emp_aff_math_biz_covid)
robust_se <- vcovHC(model4, type = "HC1")
coeftest(model4, robust_se) #hetero
summary(model4) #homo
plot(model4)
```

# Extra models with full_model1 for justification

## Model with spend_all quartiles

```{r}
# extend with spending quartiles

model_0 <- lm(revenue_all ~ wave1 + emp + spend_all_q1 + spend_all_q2 + spend_all_q3 + spend_all_q4  + engagement + merchants_all + NE + W + S, data = emp_aff_math_biz)

robust_se <- vcovHC(model_0, type = "HC1")
coeftest(model_0, robust_se) #hetero
summary(model_0) #homo
```
```{r}
vif(model_0)
```
VIF blows up when we use spend_all_q# -> why we use spend_all

## Model with different employment
```{r}
# extend with employment quartiles

model_1 <- lm(revenue_all ~ wave1 + emp_incq1 + emp_incq2 + emp_incq3 + emp_incq4 + spend_all + engagement + merchants_all + NE + W + S, data = emp_aff_math_biz)

#emp_incbelowmed + emp_incmiddle + emp_incabovemed
#emp_incq1 + emp_incq2 + emp_incq3 + emp_incq4

robust_se <- vcovHC(model_1, type = "HC1")
coeftest(model_1, robust_se) #hetero
summary(model_1) #homo
```

```{r}
vif(model_1)
```
Again high VIF


## Model without region

```{r}
model_2 <- lm(revenue_all ~ wave1 + emp + spend_all + engagement + merchants_all, data = emp_aff_math_biz)

robust_se <- vcovHC(model_2, type = "HC1")
coeftest(model_2, robust_se) #hetero
summary(model_2) #homo
```

```{r}
vif(model_2)
```

lower R^2 without region component

## Model with different engagement levels
```{r}
model_3 <- lm(revenue_all ~ wave1 + emp + spend_all + engagement_inclow + engagement_incmiddle + engagement_inchigh + merchants_all + NE + W + S, data = emp_aff_math_biz)

robust_se <- vcovHC(model_3, type = "HC1")
coeftest(model_3, robust_se) #hetero
summary(model_3) #homo
plot(model_3)
```

```{r}
vif(model_3)
```

# Original Model
```{r}
emp_aff_math_biz <- emp_aff_math_biz %>%
  filter(!is.na(revenue_all))
emp_aff_math_biz$revenue_all <- as.numeric(emp_aff_math_biz$revenue_all)
emp_aff_math_biz$merchants_all <- as.numeric(emp_aff_math_biz$merchants_all)
emp_aff_math_biz$spend_all <- as.numeric(emp_aff_math_biz$spend_all)
emp_aff_math_biz$emp <- as.numeric(emp_aff_math_biz$emp)
emp_aff_math_biz$engagement <- as.numeric(emp_aff_math_biz$engagement)

full_model1 <- lm(revenue_all ~ wave1 + emp + spend_all + engagement + merchants_all + NE + W + S, data = emp_aff_math_biz)
robust_se <- vcovHC(full_model1, type = "HC1")
coeftest(full_model1, robust_se) #hetero
summary(full_model1) #homo
plot(full_model1)
```

## FINAL MODEL
```{r}
model_4 <- lm(revenue_all ~ wave1 + emp + spend_all + engagement + merchants_professional + merchants_health + merchants_food_accommodation + merchants_other_services + merchants_retail + NE + W + S, data = emp_aff_math_biz)

robust_se <- vcovHC(model_4, type = "HC1")
coeftest(model_4, robust_se) #hetero
summary(model_4) #homo
#plot(model_4)
```

```{r}
vif(model_4)
```

